---
title: "Lab 9 - Multiple Linear Regression"
author: "Tara Jagadeesh"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use this template to follow along in Lab Week 9. Each code chunk you'll need is already created and named. 

**Lab 9 Objectives:**

- Explore multivariate data (SLO housing prices)
- Perform multiple linear regression
- Assess diagnostics
- Compare different models by AIC
- Explain model outputs
- Make a nice table of regression results
- Make predictions using a final model
- Visualize predictions with original data

###1. Load packages

- tidyverse
- stargazer

```{r packages, include = FALSE}

# a. Load packages 'tidyverse' and 'stargazer':

library(tidyverse)
library(stargazer)

```

###2. Load data (slo_homes.csv) as a df called 'homes', then filter to create a data frame called 'homes_sub' that only include homes in SLO, Arroyo Grande, Santa Maria-Orcutt, and Atascadero

```{r get_data, include = FALSE}

# a. Read in data as 'homes':

homes <- read_csv("slo_homes.csv")

# b. Filter to only include cities "San Luis Obispo", "Santa Maria-Orcutt", "Atascadero", and "Arroyo Grande", and call this new subset 'homes_sub':

homes_sub <- homes %>% 
  filter(City == "Arroyo Grande" | City == "San Luis Obispo" | City == "Atascadero" | City == "Santa Maria-Orcutt")

```

###3. Go exploring (visual) + think critically about variables

*Note: It's OK to LOOK at things separately, even if you're including all in a model together!*

Example: if we want to compare distribution of housing prices by CITY (ignoring all other variables), we can do that:

```{r by_city}

# a. Calculate mean price by city
mean_by_city <- homes_sub %>% 
  group_by(City) %>% 
  summarize(
    mean = mean(Price)
  )

# b. Visualize prices by city
by_city <- ggplot(homes_sub, aes(x = Price)) +
  geom_density(aes(color = City, fill = City), alpha = 0.3) + # Note: just to show what the geom_violin shows
  theme_classic() +
  scale_x_continuous(expand = c(0,0), limits = c(0,3e6)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Home Prices (USD)", y = "Density")

by_city

#Density distributions: bins are centered around observations (a bit different than histograms)
```

Or another question: Overall relationship between home square footage and price, separated by City? 

```{r by_sqft}

# a. Relationship between square footage and price
by_sqft <- ggplot(homes_sub, aes(x = SqFt, y = Price)) +
  geom_point(aes(color = City, pch = Status), alpha = 0.5) +
  facet_wrap(~Status)

by_sqft

# Observations here: Does relationship appear ~ linear? Anything else we can pick out re: trends, outliers, etc.? What is the general trend? Any outliers? Is there reason enough for us to omit it?

```

###4. Multiple linear regression

Multiple linear regression in R follows the same syntax we've been using so far: 

    lm(y ~ x1 + x2 + x3..., data = df_name)
    
Let's try this model a couple of different ways: 

(1) Use all available variables (saturated model) 
(2) Use only SqFt as a predictor for "home size" generally (omit Bathrooms and Bedrooms), and omit PricePerSqFt (since it's derived from two other existing variables in the model)

Use summary() to view the model output and statistics.


```{r saturated}

# a. Saturated model: 
homes_lm1 <- lm(Price ~ City + Bedrooms + Bathrooms + SqFt + PricePerSqFt + Status, data = homes_sub)

#Note: THIS IS BAD because pricepersqft is calculated from the dependent variable Price - this can lead to biased overfitting of model results 
 
# b. Model summary:

homes_lm1
# Price = -626690 - 76863(SLO) - 9068.2(Atasc) + 120160().....

#But this is a bad model..... says that as you increase bedrooms, price drops, etc etc
#SO: 1) PricePerSqFt might be throwing everything off and 2) Co-linearity -> strongly correlated variables (bedrooms, bathrooms, sqft), messes w/ coefficient terms
```

The next model: Exclude price per square foot, and bedrooms/bathrooms (since these are largely explained by square footage...)

```{r subset}

# a. Updated model with variables City, SqFt, and Status:

homes_lm2 <- lm(Price ~ City + SqFt + Status, data = homes_sub)
homes_lm2

# b. Model summary:

# Price = 106333 - 202157.5(Atasc) + 34525.4(SLO) + 

#What does this mean?
##Reference variable: Arroyo grande!!
##For SLO - if two houses are exactly the same (in terms of status and sqft), then a house would sell for 34525.4 more in SLO than Arroya Grande

#When interpreting categorical data, always do it with respect to a reference variable

#Also: for every 1 sqr foot increase in size, price goes up by $245.9 on average
#StatusRegular: A house in regular status sells for $209998.9 more than a foreclosed house
#StatusShortSale: A house in short sale status only sells for $2282.4 more than a foreclosed house (p = 0.9, so this is insignificant, which makes sense)

#However, DONT remove variables just because they are insignificant - only cut things if they conceptually dont make sense 

summary(homes_lm2)

#For multilinear regression: use adjusted R^2 (accounts for the fact that model fit goes up bc of multiple predictors)
```

Wait...but what if I wanted everything to be with respect to a Regular sale status? Then I need to change my factor levels. We've done this before, here we'll use a different function (fct_relevel) from *forcats* package in the tidyverse. 

```{r fct_relevel}

# a. Set Status to class 'factor'
class(homes_sub$Status) #See what the original class is - character
#If you feed R a character, then lm() will use it as a factor. But if you want to set the factor levels you should make sure it is a factor

homes_sub$Status <- factor(homes_sub$Status) #Make Status a factor

# b. Check to ensure it's a factor no
class(homes_sub$Status)

# c. Check levels:
levels(homes_sub$Status)
##The first one is the reference factor as default 

# d. Reassign reference level of "Status" to "Regular":
homes_sub$Status <- fct_relevel(homes_sub$Status, "Regular", "Foreclosure", "Short Sale") #Relevels in this order with Regular as the reference level

levels(homes_sub$Status)

# e. Now run the regression again - same equation, but now the reference level is different (Status = Regular): 

homes_lm3 <- lm(Price ~ City + SqFt + Status, data = homes_sub) #Now R recognizes a different reference level for status
summary(homes_lm3)


```

Interpret the coefficients and statistical outcomes above. 
Notes: 


###5. Model diagnostics

Remember, since we're concerned about *residuals* (distance that actual observations exist from model predictions), we can only evaluate some assumptions *after* running the regression. 

Then we can evaluate model diagnostics using the plot() function:

```{r diagnostics}

# a. Model diagnostics:

plot(homes_lm3)

#Overall, heteroscedasticity is okay, normality of residuals looks good, and conceptually and mathematically the model makes sense
```

Notes:

Residuals:
-Variance is equal
-Want to buy #188 -> good value

QQ:
-Normally distributed 
-Only 8/400 dont follow this 
-SO: Assumption satisfied


###6. Model comparison by Akaike Information Criterion

The AIC is a quantitative metric for model "optimization" that balances complexity with model fit. The best models are the ones that fit the data as well as possible, as simply as possible. Recall: lower AIC value indicates a *more optimal* balance - **BUT STATISTICS IS NO SUBSTITUTE FOR JUDGEMENT!!!**

```{r AIC}

# a. AIC values for each model
##Used to decide which model - BUT you should still think critically about your model

sat_aic <- AIC(homes_lm1)
sat_aic #10699.25 - means nothing on its own, need to compare it to other models

final_aic <- AIC(homes_lm3)
final_aic #11199.19

# Answer: which would you pick? 

#Based PURELY on AIC, you would pick sat_aic. However, we know that this is not the best model conceptually. 
#We should still pick homes_lm3, because: "Overall, heteroscedasticity is okay, normality of residuals looks good, and conceptually and mathematically the model makes sense""


```

###7. Regression tables with *stargazer*

```{r stargazer, results = 'asis'}

# a. Prepare a nice regression table:

lm_table <- stargazer(homes_lm1, homes_lm3, type = "html")
#Have to knit to html to see it

# Note: If you want to work with this in Word, save to html, open, copy and paste into Word. 

```

###8. Making predictions

Using your final selected model, predict the housing price for a range of home sizes, sale status, and city. 

The predict() function uses the following syntax:

      predict(model_name, newdata = new_data_name)
      
Defaults are to exclude the prediction SE and mean confidence interval - if you want to include, use arguments

      se.fit = TRUE
      interval = "confidence" 
      interval = "prediction"

First, you need to create a new data frame of values that contain ALL NECESSARY VARIABLES **with the same variable names AND level strings**.

```{r df_new}

# First, make a new data frame

# Note that the df_new created below has the SAME variable names and level strings as the original model data (otherwise R won't know how to use it...)
# Work through this on your own to figure out what it actually does:

df_new <- data.frame(City = rep(c("San Luis Obispo",
                                  "Santa Maria-Orcutt",
                                  "Atascadero",
                                  "Arroyo Grande"), 
                                each = 60), 
                     SqFt = rep(seq(from = 500,
                                    to = 3500, 
                                    length = 20), 
                                times = 12), 
                     Status = rep(c("Regular",
                                    "Foreclosure",
                                    "Short Sale"), 
                                  times = 12, 
                                  each = 20))

```

Make predictions for the new data using predict():

```{r predict}

# a. Make predictions using the new data frame:

price_predict <- predict(homes_lm3, 
                         newdata = df_new, 
                         se.fit = TRUE, 
                         interval = "confidence")

# b. Bind predictions to the data to make it actually useful:

predict_df <- data.frame(df_new, price_predict)

```

Then visualize it!

```{r graph, echo = FALSE}

# Create a line graph with square footage on the x-axis, predicted price on y-axis, with color dependent on City and facet_wrapped by sale status (follow along):

predict_graph <- ggplot(predict_df, aes(x = SqFt, y = fit.fit))+
  geom_line(aes(color = City))+
  facet_wrap(~Status)
predict_graph

```

END LAB